Data Specialization for a Staged Lambda Calculus

### 2015-02-27 paper/semantics.tex

\fbox{\parbox{5in}{
\paragraph{Blend.}
the
function \texttt{qs} does \emph{not} depend on the rank \texttt{k};
for example, the list is partitioned before \texttt{k} is ever
used. Moreover, while the recursive calls to \texttt{qSelect} are
guarded by a comparison to \texttt{k}, \texttt{k} does not determine
the arguments to those recursive calls---it only determines which call
is made. Thus, if we are willing to change the evaluation order of the
language (and evaluate under the \texttt{case} statement), it should
be possible to force \emph{all} the computations involving \texttt{l}
to be performed before \emph{any} of the computations involving
\texttt{k}.

The \texttt{case} expression in \texttt{qSelect} depends on \texttt{k}, whose
type $\fut\rmint$ indicates that it is only available to second-stage
computations. Since \texttt{qsStaged} itself produces a $\fut\rmint$, the
remainder of the function is second-stage code.

}}

@@

>>>>

@@

\crem{blend in the rest of the text in this section:}
%
\texttt{qsStaged} has a more precise type---it is a first-stage function which
takes a $\curr\mathrm{list}$ (an integer list now) and a $\fut\rmint$ (an
$\rmint$ \emph{in the future}), and returns a $\fut\rmint$.

Indeed, we case on the list as usual, and in the \texttt{Cons} branch,
\texttt{part}ition it. The \texttt{case} expression in \texttt{qSelect} depends
on \texttt{k}, whose type $\fut\rmint$ indicates that it is only available to
second-stage computations. Since \texttt{qsStaged} itself produces a
$\fut\rmint$, the remainder of the function is second-stage code.

\crem{does this remark go anywhere?}

Thus, if we are willing to change the evaluation order of the
language (and evaluate under the \texttt{case} statement), it should
be possible to force \emph{all} the computations involving \texttt{l}
to be performed before \emph{any} of the computations involving
\texttt{k}.
